Capturing the geometric structure of our experiences and how we remember them

### 2018-07-25 main.tex
intro and results edits

\section{Abstract}
{
The human memory system is adept at cataloging the rich dynamics of ongoing experience. A hallmark feature of human experiences is that they unfold in a structured and predictable manner, governed by the laws of the physical systems that produce them. However, this temporal structure is typically absent (or severly impoverished) in traditional trial-based memory experiments. As a consequence, our understanding of how the inherent temporal structure of our life experiences influences memorability and memory organization is limited. Using a novel computational modelling approach, we investigate how people verbally recall continuous videos by characterizing and relating the thematic dynamics, or ``trajectories,'' of the stimulus and participants' recalls. We found that despite large differences in the quantity of information recalled across participants, the overall shape of the stimulus was recapitulated by most of the participants. These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.

@@

Human memory studies typically assess memory for the individual contents of an experience, such as the people, places and things/items encountered. However, real life experiences can also be described in terms of their temporal structure (i.e. how the contents relate over time).  Naturalistic experiences are typically autocorrelated in space and time on short timescales: the contents of an experience are often similar from moment to moment and change gradually. For example, many of us spend our mornings/evenings at home and much of our days at work (e.g. an office). While the specific happenings within these contexts differ, they are often highly correlated (and thus predictable) over time. Furthermore, our experiences can also be characterized by longer timescale correlations. For example, consider returning to your office after taking a lunch break. The spatial contexts before and after lunch are highly similar, but separated in time. Thus, while evaluating the ability to recall specific contents of an experience is one way to assess memory, another important dimension to consider is the temporal structure of the remembered experience, and how the temporal structure of the memory relates to the original experience. However, these recurrent and gradually changing temporal dynamics are not typically present in traditional memory studies, but are likely critical if we wish to understand how our memory systems remember our life experiences.

@@

\subsubsection{A model to capture the temporal and semantic structure of naturalistic stimuli}
We fit a topic model \citep{BleiEtal03} to (manually annotated) text descriptions of scenes from the video. The text descriptions contained details of the scene such as the characters, location, and a short summary of the scene (see Fig.\ref{fig:schematic} for example text, see \ref{sec:methods} for analysis details). We then transformed the text descriptions using the (same) topic model, resulting in a scenes (1976) by topics (100) matrix, where each row of the matrix represents a probabalistic mixture of topics discovered in that scene (See Fig.\ref{fig:schematic} for example topic vectors). As depicted in Fig.~\ref{fig:model}a, the topic vectors are sparse and change slowly over time. Furthermore, there are clear transitions from one topic `state' to the next, possibly indexing scene transitions in the stimulus. To get a better handle on this temporal structure, we computed a timepoint (1976) by timepoint (1976) correlation matrix of the video model (Fig.~\ref{fig:model}c).  This correlation matrix reveals that the model has a strong, block-diagonal structure. Another noteworthy characteristic is that there is very little correlation between blocks (i.e. the off-diagonal values are small), suggesting the representations of each block are unique and highly descriminable.

\subsubsection{Modeling verbal recall}
After watching the video, participants verbally recalled (in order) as much of the episode as they could.  We used the same topic model (fit with the text descriptions of the video) to transform participants' verbal recall transcripts. The result was a sentences (range: 68 to 294, mean: 131.94) by topics (100) matrix for each participant, where each row represents a probabalistic mixture of topics for a given window of sentences during recall. An example participant's (\#13) recall model is plotted in Fig.~\ref{fig:model}c. Like the video model, topic vectors were sparse and changed gradually.  Note that the topics were derived solely from the text descriptions of the video, and so the topic models estimated for recall are dependent on the topics present in the video. This approach effectively projects the participants' verbal recall into the video's `topic space'', allowing us to directly compare them to the video model.

Next, we investigated the temporal structure of the recall matrices. For each participant, we computed a timepoint (68 to 294) by timepoint (68 to 294) correlation matrix from the recall models. An example participant's (\#13) correlation matrix is plotted in Fig.~\ref{fig:model}d). Like the video model, each participant's recall correlation matrix exhibited a strong block-diagonal structure (Supp Fig.~\ref{fig:corrmats}). Notably, this suggests that participants recounted the video in discriminable segments, potentially indexing the recall of specific events from the video.

\subsubsection{Segmenting the video and recall models into `events'}
As described above, a striking feature of the video and recall correlation matrices (see Fig.~\ref{fig:model}) is a strong, block structure along the diagonal of the matrices.  We hypothesized that this structure might arise from transient stability in the contents of the video (i.e. ``events''), as well as the subsequent language used to recount each event of the video. If true, we reasoned that it would be possible to identify the specific event a participant is describing by comparing the topic vector for a particular recall ``event'' with the topic vector for each video event. In other words, the video event that is most similar (i.e. correlated) to a particular recall event is the event that the participant is most likely describing.

To test this idea, first we segmented the video and recall models in $k$ events (i.e. states) using a hidden Markov model (~\citep{BaldEtal17}). The algorithm determined 34 events for the video model and a range of values (range: 8-27; mean: 15.41; SD: 5.6) for the recall models (see ~\ref{sec:methods} for details on choosing an optimal $k$ value).  The events discovered for the model and a representative participant (\#13) are highlighted as yellow rectangles outlining blocks along the diagonal of the correlation matrices (Fig.~\ref{fig:model}b,d).

Next, we created a video event model by averaging together neighboring topic vectors that the were classified to be in the same event, resulting in an events (34) by topics (100) matrix (Fig.~\ref{fig:model}e).  We performed the same procedure for the recall matrices (see Fig.~\ref{fig:model}f for example). Then, we computed the correlation between video and recall event models, resulting in a video events (34) by recall events (8-27, depending on the participant) correlation matrix (Fig/~\ref{fig:model}g). These matrices represent the similarity (correlation) between each video event and each recall event (for each participant). To determine which video event a particular recall event was most likely describing, we computed the index of the video event with the highest correlation to the recall event (i.e. the argmax).  This is depicted in Fig.~\ref{fig:model}g as the cells highlighted with a yellow border. Interestingly, our algorithm suggests that the example participant recalled the events largely in order, occasionally skipping a video event.

Then, we computed a group-averaged recall event model and video-recall event correlation matrix. For each participant (and each recall event), we sorted the recall event vectors (across all participants) according to the video event that they were most highly correlated to. We then averaged the recall event vectors within each group. This yielded an average recall event vector for all but one (of 34) video events, since no participant remembered one of the events according to our model. Lastly, we computed an average recall event (34) by video event (34) correlation matrix, and highlighted the cell with the highest correlation value with a yellow border (Fig.~\ref{fig:model}h). Notably, this matrix displayed high correlation values along the diagonal and low correlations in the off-diagonal cells. This suggests that on average, participants were able to recapitulate the events in the episode in a specific, highly descriminable way.
% NOTE: ^ possibly need a null model for the group analysis.  I think it could be to get a result that looks like this even if we scrambled the text. My concern is that I'm grouping the recall events by the video event they were most similar to, and then averaging, which may give rise to a smooth looking plot without any real information there...

@@

\subsubsection{Characterizing naturalistic memory with traditional list-learning analyses}
Just like in a traditional `free-recall' list-learning experiment where participants view a list of words and then verbally recall them, our video-recall matching analysis approach affords us the ability to analyze memory in the same way. The recalled events can be treated as ``items'' analogous to words recalled in a list-learning study. In our first set of analyses, we sought to characterize memory performance/dynamics by extending classic analyses originally designed for list-learning experiments to more naturalistic settings.

First, we asked whether the estimated number of recall events ($k$) by participant was related to hand-annotated accuracy as published in Chen et al. (2017).  We found a strong positive correlation where subjects with a greater number of recall events also had better overall memory performance (Pearson's $r$(16)=.67, $p$=.003). Then, we considered how participants initiated the recall sequence (known in the literature as the `probability of first recall' or `PFR'). This metric quantifies the likelihood of initiating a sequence of recalls based on an item's serial position during encoding. We found that participants tended to initiate their recall sequences with the first few events (Fig.~\ref{fig:list-learning}a), which is qualitatively very similar to previously published list learning experiments (REF). Next, we considered another well-studied memory measure in the list-learning literature, the lag conditional response probability curve (or lag-CRP) ~\citep{Kaha96}. Given the recall of a stimulus in study position $n$, the lag-CRP quantifies the probability that the next item recalled will be of lag $i$. The result suggests a strong bias to transition sequentially events in the forward direction Fig.~\ref{fig:list-learning}b. This can be seen by the high probability values in the lag ~1-5 positions compared to the rest of the transition possibilties. Notably, participants were instructed to recall as much as they could in order, and the shape of this lag-CRP is qualitatively similar to list-learning studies of serial recall (REF), where participants are required to recall lists of words in their serial order. Finally, we assessed memory performance for each event in the video as a function of its serial position during encoding (Fig.~\ref{fig:list-learning}c). We found that there was substantial variability in memory for the different events (STAT?). Furthermore, the shape of the curve is different than what is typically found in list learning experiments (we'll come back to this point in the dicussion).

 We also considered two additional across-participant measures of recall that characterize memory organization: temporal clustering and semantic clustering. Temporal clustering (REF?) measures the extent to which participants tend to cluster recalls of events encoded nearby in time.  For example, if a participant recalled in exact serial order, the score would be 1 whereas if they sequentially recalled randomly positioned items the score would be ~.5.  We found that participants who clustered in time also recalled a greater number of events (Pearson's $r$(16)=.62, $p$=.007). Next, we assessed semantic clustering.  Similar to temporal clustering, semantic clustering measures the extent to which participants cluster their responses by semantic similarity.  We used the topic vector representations of each event as a proxy for semantic information contained in that event.  For each transition, we computed the rank percentile of the of the correlation between the the current event and then next event. For example, if a participant sequentially recalled events where each successive recall was the event that was most semantically similar to the previous, the semantic clustering score would be 1.  If the events were recalled in a random order with respect to semantic information, the clustering score would be ~.5. We found that the semantic clustering score was related to memory performance across participants (Pearson's $r$(16)=.55, $p$=.02).  Thus, participants who organized their recalls with respect to the semantic information contained in the scene had better memory performance.

 \subsubsection{Measuring the quality of naturalistic memories}
 Representing the video and verbal recall as events in `topic space' also affords us the ability to characterize the quality of recall in a more fine-grained and nuanced way than what was previously possible. To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the movie model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered. Then, we computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events our of order or with low precision the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants ($p$<.05; participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to memory performance across participants (Pearson's $r$(16)=-.57, $p$=.016). Thus, a closer match between the video and recall event models was related to better recall performance.

 Next, we tested whether participants who recalled more events were also more precise in their recollections. For each participant, we computed the correlation between each recall event and its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event indexing how similar the recall event was to its matching movie event (i.e the `precision' of the recall). We then averaged the correlations within participant. In line with our prediction, there was a strong correlation between memory performance and precision suggesting that participants who remembered more events also remembered them more veridically (Pearson's $r$(16)=.74, p=.0006). Next, we considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e. 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher memory performance also had higher distinctiveness scores (Pearson's $r$(16)=.8, p=.0001).

 Lastly, we tested whether participants with better memory performance were also more likely to remember the events in order.  For each participant, we computed the Spearman rank correlation between the order of events that the participant recalled and the actual order of events (filtering events that were actually recalled).  We found that participants who recalled more events also recalled more of them in order (Pearson's $r$(16)=.5, p=.04). In summary, we found that better memory performance was associated with more precise and distinctive recall.

\subsubsection{The fundamental shape of an experience is preserved in recall}
The analyses described in the previous sections are useful in that they allow us to quantify memory dynamics for naturalistic stimuli in a way that is comparable to a vast literature on trial-based list-learning experiments. However, these methods fail to capture the rich temporal structure present in naturalistic stimuli and associated recall of those experiences. Thus, our next set of analyses test whether the ``shape'' of an experience is preserved in memory. We hypothesized that despite substantial individual variability in the precision and amount of content recalled across by participants, the overall shape of the episode would be recapitulated by the majority of participants.

@@

To visualize the relationship between the video and recall event models, we embedded them into a 2D space (using the UMAP dimensionality reduction algorithm, for details see \ref{sec:methods}) where the points represent video/recall events and the distance between them represents their similarity in ``topic space'' (Fig.~\ref{fig:trajectory}a).   We observed that the two models have a very similar temporal evolution and geometric structure. To further quantify this correspondence and to characterize how participants navigated through the space during recall, we analyzed the angle between successive recall events. We created a grid of evenly spaced points (.25 units) in the 2D ``topic space''. For each point on the grid, we drew a circle (radius=.25) around the point and grouped together all recall transitions that intersected the circle (across subjects). To visualize the average angle, we converted each transition angle to a unit vector and then averaged the vectors together. To assess consistency in the direction of the recall transition across participants, we performed a Rayleigh test (p<.001, corrected at p<.05 using permutation procedure described in \ref{sec:methods}). We found that on average, participants' recalls flowed along the same path as the video model (Fig.~\ref{fig:trajectory}b). Each participant's recall path can be seen in Fig.~\ref{fig:trajectory}c.

@@

\subsubsection{Memorability by event and topics}
Another advantage of representing naturalistic stimuli/memory using this approach is that the event models can be mapped back to the language used to fit the model(s).  In this next analysis, we 1) quantified the memorability of each video event and 2) used memorability to plot the top words for the most and least memorable scenes, and across the entire video.

To measure video event memorability, we computed the correlational distance between each video event and the closest recall event and then averaged those values within participant. We used this measure of memory to create a weighted average over all video events, where video events with better memory were weighted more heavily. Then, we extracted the top 200 words and created ``wordles'' from the most memorable scenes of the video (see Methods for details). The result reveal that scenes containing "Sherlock" and "John" are highly memorable, and also "medium", "street" and "baker" (Fig.~\ref{fig:topics}b). To find words related to the least memorable scenes, we inverted the "memorability" weights and extracted the top 200 words.  Scenes containing "Mike" and "Molly" were least remembered, and also words like "conference", "hospital" and "psychotherapist" (Fig.~\ref{fig:topics}b).

Then, we took a closer look at the themes contained in the top/bottom 3 events. We used the "memorability" weights to select the top/bottom events. We performed the same analysis as described above, but separately for the video event and average recall event vectors representing each of the top/bottom 3 remembered events. In Figure~\ref{fig:topics}c, the trajectory represents a low-dimensional embedding of the video model and the large colored dots represent the video events. The smaller colored dots represent the distribution of individual recall events across all participants, colored by the closest video event. We plotted wordles for the top 3 (circled in green) and bottom 3 (red) most/least remembered events, where the video is represented in the left half of the circle and the average recall is represented in the right half of the circle. Visual inspection reveals that the words contained in the memorability weighted worldles (Fig.~\ref{fig:topics}b) overlap with the individual top/bottom event wordles (Fig.~\ref{fig:topics}c). These analyses uncover the contents of naturalistic stimuli and accompanying memories, and highlight the flexibility in transforming between human and machine readable representations of the data afforded by our modeling approach.

@@

\subsubsection{Brain regions that mirror the correlational structure of naturalistic memories}.
While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event by event similarity structure (e.g. the off-diagonal values of the recall correlation matrices) differed by individual (Supp. Fig.~\ref{fig:corrmats}). In the final analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls.

The amount of time/number of sentences used describing each scene differed drastically between participants and as compared to the original video \cite{ChenEtal17}. To account for these temporal non-linearities between the video and recall, we used dynamic time warping to align each participant's recall model and the movie model (DTW ref).  We used the resulting alignment path to warp the fMRI timeseries (for each voxel) as well as the recall model timeseries.  Then using a searchlight procedure (5mm), we computed the timepoint-by-timepoint neural correlation matrix within each searchlight. To measure the similarity between the neural and behavioral correlation matrices, we vectorized and then correlated them. The result was a correlation value for each voxel representing the extent to which the brain data mirrored the recall data.  Finally, to test the significance of the correlation values, we phase shifted the recall models (100 permuations) and recomputed the correlation values. Significance was determined by ``real'' correlation values that exceeded the 95th percentile of this null distribution. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants'' recalls.

>>>>

\section{Abstract}
{
The human memory system is adept at cataloging the rich dynamics of ongoing experience. A defining feature of our experiences is that they unfold in a structured and predictable manner, governed by the laws of the physical systems that produce them. However, this temporal structure is typically absent (or severly impoverished) in traditional trial-based memory experiments. As a consequence, our understanding of how the inherent temporal structure of our life experiences influences memorability and memory organization is limited. Using a novel computational modelling approach, we investigate how people verbally recall continuous videos by characterizing and relating the thematic dynamics, or ``trajectories,'' of the stimulus and participants' recalls. We found that despite large differences in the quantity of information recalled across participants, the overall shape of the stimulus was recapitulated by most of the participants. These findings provide a window into which aspects of naturalistic experiences must be preserved, and which might be more flexible, in considering whether and how those experiences are remembered.

@@

Human memory studies typically assess memory for the ``items'' within an experience, such as the people, places and things encountered. However, real life experiences can also be described in terms of their temporal structure (i.e. how the contents relate/change over time).  Naturalistic experiences are typically autocorrelated in space and time on short timescales: the contents of an experience are often similar from moment to moment and change gradually. For example, many of us spend our mornings/evenings at home and much of our days at work (e.g. an office). While the specific happenings within these contexts differ, they are often highly correlated (and thus predictable) over time. Furthermore, our experiences can also be characterized by longer timescale correlations. For example, consider returning to your office after taking a lunch break. The spatial contexts before and after lunch are highly similar, but separated in time. Thus, while evaluating the ability to recall specific contents of an experience is one way to assess memory, another important dimension to consider is the temporal structure of the remembered experience, and how the temporal structure of the memory relates to the original experience. However, these recurrent and gradually changing temporal dynamics are not typically present in traditional memory studies, but are likely critical if we wish to understand how our memory systems remember our life experiences.

@@

\subsubsection{A model to capture the temporal and semantic structure of naturalistic stimuli}
We fit a topic model \citep{BleiEtal03} to hand-annotated text descriptions of scenes from the video. The text descriptions contained details of the scene such as the characters, location, and a short summary of the scene (see Fig.\ref{fig:schematic} for example text, see \ref{sec:methods} for analysis details). We then transformed the text descriptions using the (same) topic model, resulting in a scenes (1976) by topics (100) matrix, where each row of the matrix represents a probabalistic mixture of topics discovered in that scene (See Fig.~\ref{fig:schematic} for example topic vectors). As depicted in Fig.~\ref{fig:model}a, the topic vectors are sparse and change slowly over time. Furthermore, there are clear transitions from one topic `state' to the next, possibly indexing scene transitions in the stimulus. To get a better handle on this temporal structure, we computed a timepoints (1976) by timepoints (1976) correlation matrix of the video model (Fig.~\ref{fig:model}c).  This correlation matrix reveals that the model has a strong, block-diagonal structure. Another noteworthy characteristic is that there is very little correlation between blocks (i.e. the off-diagonal values are small), suggesting the representations of each block are unique and highly descriminable.

\subsubsection{Modeling verbal recall}
After watching the video, participants verbally recalled (in order) as much of the episode as they could.  We used the same topic model (fit with the text descriptions of the video) to transform participants' verbal recall transcripts. The result was a sentences (range: 68 to 294, mean: 131.94) by topics (100) matrix for each participant, where each row represents a probabalistic mixture of topics for a given window of sentences during recall. An example participant's (\#13) recall model is plotted in Fig.~\ref{fig:model}b. Like the video model, topic vectors were sparse and changed gradually.  Note that the topics were derived solely from the text descriptions of the video, and so the topic models estimated for recall are dependent on the topics present in the video. This approach effectively projects the participants' verbal recall into the video's ``topic space'', allowing us to directly compare them to the video model.

Next, we investigated the temporal structure of the recall matrices. For each participant, we computed a timepoint (68 to 294) by timepoints by timepoints (range: 68 to 294) correlation matrix from the recall models. An example participant's (\#13) correlation matrix is plotted in Fig.~\ref{fig:model}d). Like the video model, each participants' recall correlation matrix exhibited a strong block-diagonal structure (Supp Fig. 1). Notably, this suggests that participants recounted the video in discriminable segments, potentially related to the recall of specific events (or scenes) from the video.

\subsubsection{Segmenting the video and recall models into `events'}
As described above, a striking feature of the video and recall correlation matrices is a strong, block structure along the diagonal of the matrices (see Fig.~\ref{fig:model}c,d).  We hypothesized that this structure might arise from transient stability in the contents of the video (i.e. ``events''), as well as the subsequent language used to recount each event of the video. If true, we reasoned that it would be possible to identify the specific event a participant is describing by comparing the topic vector for a particular recall ``event'' with the topic vector for each video event. In other words, the video event that is most similar (i.e. correlated) to a particular recall event is the event that the participant is most likely describing.

To test this idea, first we segmented the video and recall models in $k$ events (i.e. states) using a hidden Markov model ~\citep{BaldEtal17}. Our algorithm determined 34 events for the video model and a range of values (range: 8-27; mean: 15.41; SD: 5.6) for the recall models (see Methods for details on choosing an optimal $k$ value).  The events discovered for the model and a representative participant (\#13) are highlighted as yellow rectangles outlining blocks along the diagonal of the correlation matrices (Fig.~\ref{fig:model}b,d).

Next, we created a video ``event model by'' averaging together neighboring topic vectors that the were classified to be in the same event, resulting in an events (34) by topics (100) matrix (Fig.~\ref{fig:model}e).  We performed the same procedure for the recall matrices (see Fig.~\ref{fig:model}f for example). Then, we computed the correlation between video and recall event models, resulting in a video events (34) by recall events (8-27, depending on the participant) correlation matrix (Fig~\ref{fig:model}g). These matrices represent the similarity (correlation) between each video event and each recall event (for each participant). To determine which video event a particular recall event was most likely describing, we computed the index of the video event with the highest correlation to the recall event (i.e. the argmax).  This is depicted in Fig.~\ref{fig:model}g as the cells highlighted with a yellow border. Interestingly, our algorithm suggests that the example participant recalled most of the video events in order.

Then, we computed a group-averaged recall event model and video-recall event correlation matrix. For each participant (and each recall event), we sorted the recall event vectors (across all participants) according to the video event with the highest correlation. We then averaged the recall event vectors within each group. This yielded an average recall event vector for all but one (of 34) video events, since no participant remembered one of the events according to our model. Lastly, we computed an average recall event (34) by video event (34) correlation matrix, and highlighted the cell with the highest correlation value with a yellow border (Fig.~\ref{fig:model}h). Notably, this matrix displayed high correlation values along the diagonal and low correlations in the off-diagonal cells. This suggests that on average, participants were able to recapitulate the events in the episode in a specific, highly descriminable way.
% NOTE: ^ possibly need a null model for the group analysis.  I think it could be to get a result that looks like this even if we scrambled the text. My concern is that I'm grouping the recall events by the video event they were most similar to, and then averaging, which may give rise to a smooth looking plot without any real information there...

@@

\subsubsection{Characterizing naturalistic memory with traditional list-learning analyses}
Just like in a traditional ``free-recall'' list-learning experiment where participants view a list of words and then verbally recall them, our video-recall matching analysis approach affords us the ability to analyze memory in the same way. The recalled events can be treated as ``items'' analogous to words recalled in a list-learning study. In our first set of analyses, we sought to characterize memory performance/dynamics by extending classic analyses originally designed for list-learning experiments to more naturalistic settings.

First, we asked whether the estimated number of recall events ($k$) by participant was related to hand-annotated accuracy as published in Chen et al. (2017).  We found a strong positive correlation where subjects with a greater number of recall events also had better overall memory performance (Pearson's $r$(16)=.67, $p$=.003). Then, we considered how participants initiated the recall sequence (known in the literature as the `probability of first recall' or `PFR'). This metric quantifies the likelihood of initiating a sequence of recalls based on an item's serial position during encoding. We found that participants tended to initiate their recall sequences with the first few events (Fig.~\ref{fig:list-learning}a), which is qualitatively very similar to previously published list learning experiments (REF?). Next, we considered another well-studied memory measure in the list-learning literature, the lag conditional response probability curve (or lag-CRP) ~\citep{Kaha96}. Given the recall of a stimulus in study position $n$, the lag-CRP quantifies the probability that the next item recalled will be of lag $i$. The result suggests a strong bias to transition sequentially events in the forward direction (Fig.~\ref{fig:list-learning}b). This can be seen by the high probability values in the lag ~1-5 positions compared to the rest of the transition possibilties. Notably, participants were instructed to recall as much as they could in order, and the shape of this lag-CRP is qualitatively similar to list-learning studies of serial recall (REF?), where participants are required to recall lists of words in their serial order. Finally, we assessed memory performance for each event in the video as a function of its serial position during encoding (Fig.~\ref{fig:list-learning}c). We found that there was substantial variability in memory for the different events (STAT). Furthermore, the shape of the curve is different than what is typically found in list learning experiments.

 We also considered two additional across-participant measures of recall that characterize memory organization: temporal clustering and semantic clustering. Temporal clustering (REF?) measures the extent to which participants tend to cluster recalls of events encoded nearby in time.  For example, if a participant recalled in exact serial order, the score would be 1 whereas if they sequentially recalled randomly positioned items the score would be \~.5.  We found that participants who clustered in time also recalled a greater number of events (Pearson's $r$(16)=.62, $p$=.007). Next, we assessed semantic clustering.  Similar to temporal clustering, semantic clustering measures the extent to which participants cluster their responses by semantic similarity.  We used the topic vector representations of each event as a proxy for semantic information contained in that event.  For each transition, we computed the rank percentile of the of the correlation between the the current event and then next event. For example, if a participant sequentially recalled events where each successive recall was the event that was most semantically similar to the previous, the semantic clustering score would be 1.  If the events were recalled in a random order with respect to semantic information, the clustering score would be ~.5. We found that the semantic clustering score was related to memory performance across participants (Pearson's $r$(16)=.55, $p$=.02).  Thus, participants who organized their recalls with respect to the semantic information contained in the scene had better memory performance.

 \subsubsection{Measuring the quality of naturalistic memories}
 Representing the video and verbal recall as events in ``topic space'' also affords us the ability to characterize the quality of recall in a more fine-grained and nuanced way than what was previously possible. To quantify the similarity between the video model and individual recall models, we considered a number of novel metrics.  First, we tested whether each participant's recall model matched the movie model in a general sense. To do this, for each participant we filtered the video model to only include the events that the participant remembered. Then, we computed the root mean squared difference (RMSD) between the video model and the recall model. As an example, if the participant remembered all the events in order (with perfect precision), the expected distance value would be 0. However, if they remembered a subset of events, events our of order or with low precision the expected distance would be greater than 0. To assess significance, we performed a permutation test where we circularly shifted the recall model (10000 times) and recomputed the RMSD. The recall model significantly matched the video model for nine of the participants ($p$<.05; participants: 3-4, 8-13, 17 and the p-value for the rest of the participants was less than .25). Furthermore, the RMSD values were significantly correlated to hand annotated memory performance across participants (Pearson's $r$(16)=-.57, $p$=.016). Thus, a closer match between the video and recall event models was related to better recall performance.

 Next, we tested whether participants who recalled more events were also more precise in their recollections. For each participant, we computed the correlation between each recall event and its matching video event (only for the events which they recalled). This resulted in a single number for each recalled event indexing how similar the recall event was to its matching movie event (i.e the ``precision'' of the recall). We then averaged the correlations within participant. In line with our prediction, there was a strong correlation between hand annotated memory performance and precision suggesting that participants who remembered more events also remembered them more veridically (Pearson's $r$(16)=.74, p=.0006). Next, we considered the distinctiveness of each recall event. That is, how uniquely a recall event matched a given video event compared to all other video events. We hypothesized that participants with high memory performance might describe each event in a more distinctive way (relative to those with lower memory performance who might describe events in a more general way). To this end, we computed a `distinctiveness' score for each participant (i.e. 1 - the correlation between a recall event and all non-matching video events).  Then, we averaged this measure over recall events within participant.  We found that participants with higher hand annotated memory performance also had higher distinctiveness scores (Pearson's $r$(16)=.8, p=.0001).

 Lastly, we tested whether participants with better memory performance were also more likely to remember the events in order.  For each participant, we computed the Spearman rank correlation between the order of events that the participant recalled and the actual order of events (filtering events that were actually recalled).  We found that participants who recalled more events also recalled more of them in order (Pearson's $r$(16)=.5, p=.04). In summary, we found that better memory performance was associated with more precise, distinctive and ordered recalls.

\subsubsection{The shape of a naturalistic experience is preserved in recall}
The analyses described in the previous sections are useful in that they allow us to quantify memory dynamics for naturalistic stimuli in a way that is comparable to a vast literature on trial-based list-learning experiments. Furthermore, we introduce novel measures (precision and distinctiveness) which provide additional details regarding the veracity of naturalistic memories. However, these methods fail to capture the rich temporal structure present in naturalistic stimuli and associated recall of those experiences. Thus, our next set of analyses test whether the ``shape'' of an experience is preserved in memory. We hypothesized that despite substantial individual variability in the precision and amount of content recalled across by participants, the overall shape of the episode would be recapitulated by the majority of participants.

@@

To visualize the relationship between the video and recall event models, we embedded them into a 2D space (using the UMAP dimensionality reduction algorithm, for details see \ref{sec:methods}) where the points represent video/recall events and the distance between them represents their similarity in ``topic space'' (Fig.~\ref{fig:trajectory}a).   We observed that the two models have a very similar temporal evolution and geometric structure. To further quantify this correspondence and to characterize how participants navigated through the space during recall, we analyzed the angle between successive recall events. We created a grid of evenly spaced points (.25 units) in the 2D ``topic space''. For each point on the grid, we drew a circle (radius=.25) around the point and grouped together all recall transitions that intersected the circle (across subjects). To visualize the average angle, we converted each transition angle to a unit vector and then averaged the vectors together. To assess consistency in the direction of the recall transition across participants, we performed a Rayleigh test (p$<$.001, corrected at p$<$.05 using permutation procedure described in Methods).Thus participants' recalls followed the same path as the video model (Fig.~\ref{fig:trajectory}b), suggesting the shape of the stimulus was preserved despite idiosyncratic differences between participants in their recalls.

@@

\subsubsection{Memorability by event and topics}
A noteworthy advantage of representing naturalistic stimuli/memory using this approach is that the event models can be mapped back to the language used to fit the model(s).  This allows for analysis and visualization of the language used when participant's recounted their experiences.  In this next analysis, we 1) quantified the memorability of each video event and 2) used memorability to plot the top words for the most and least memorable scenes, and also aggregated across the entire video.

To measure video event memorability, we computed the correlation between each video event and the closest recall event and then averaged those values within participant (Fig.~\ref{fig:topics}a). We used event memorability to create a weighted average over all video event vectors, where video events with better memory were weighted more heavily. Then, we extracted the top 200 words and created ``wordles'' representing the most memorable themes in the video. The result reveal that scenes containing ``Sherlock'' and ``John'' were highly memorable, and also scenes containing words such as ``medium'', ``street'' and ``baker'' (Fig.~\ref{fig:topics}b). To find words related to the least memorable scenes, we inverted the memorability weights and again extracted the top 200 words.  Scenes containing ``Mike'' and ``Molly'' were least remembered, and also words like ``conference'', ``hospital'' and ``psychotherapist'' (Fig.~\ref{fig:topics}b).

Then, we took a closer look at the themes contained in the top/bottom 3 events, using the memorability weights to select the top/bottom events. We performed the same analysis as described above, but separately for the video and average recall event vectors representing each of the top/bottom 3 remembered events. In Figure~\ref{fig:topics}c, the trajectory represents a low-dimensional embedding of the video model and the large colored dots represent the video events. The smaller colored dots represent the distribution of individual recall events across all participants, colored by the closest video event. We plotted wordles for the top 3 (circled in green) and bottom 3 (red) most/least remembered events, where the video is represented in the left half of the circle and the average recall is represented in the right half of the circle. Visual inspection reveals that the words contained in the memorability weighted worldles (Fig.~\ref{fig:topics}b) overlap with the individual top/bottom event wordles (Fig.~\ref{fig:topics}c). These analyses shed light on the contents of naturalistic stimuli and accompanying memories, and highlight the flexibility in transforming between human and machine readable representations of the data afforded by our modeling approach.

@@

\subsubsection{Brain regions that mirror the correlational structure of naturalistic memories}.
While each participant remembered roughly the same sequence of events, there were subtle idiosyncracies in the way each individual described the events. For example, the event by event similarity structure (e.g. the off-diagonal values of the recall correlation matrices) differed across individuals (Supp. Fig.~\ref{fig:corrmats}). We hypothesized that idiosyncracies in the recall dynamics across subjects might be related to the temporal structure of brain networks are representing the events during video viewing. Thus in the last analysis, we identified brain areas whose responses at the time of viewing exhibited the same (segmented) idiosyncratic dynamics that later appeared in that participants’ recalls.

The amount of time/number of sentences used describing each scene differed drastically between participants and as compared to the original video \cite{ChenEtal17}. To account for these temporal non-linearities between the video and recall, we used dynamic time warping to align each participant's recall model and the movie model (DTW REF).  We used the resulting alignment path to warp the fMRI timeseries (for each voxel) as well as the recall model timeseries.  Then using a searchlight procedure (5mm), we computed the timepoint-by-timepoint neural correlation matrix within each searchlight. To measure the similarity between the neural and behavioral correlation matrices, we vectorized and then correlated them. The result was a correlation value for each voxel representing the extent to which the brain data mirrored the recall data.  Finally, to test the significance of the correlation values, we phase shifted the recall models (100 permuations) and recomputed the correlation values. Significance was determined by ``real'' correlation values that exceeded the 95th percentile of this null distribution. The resulting map suggests that the correlational structure of a network of brain regions including the ventromedial prefrontal cortex, anterior cingulate, and right medial temporal lobe mirror the idiosyncratic structure of participants'' recalls.
